---
title: "abby_Rassign"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("../abby_546X/UNIX_Assignment")
library(tidyverse)
```

I will also note that the working directory of the terminal must also be the UNIX_Assignment folder.

## Data Inspection

This bit of code reads in the fang genotypes file from the UNIX assignment.
```{r}
fang <- read.delim('fang_et_al_genotypes.txt', header = TRUE, sep= "\t")
```

Next I read in the snp position file.
```{r}
snp_position <- read.delim('snp_position.txt', header = TRUE, sep= "\t")
```

To inspect these files I tried several different commands.

The following provides all of the columm names for both data frames.
```{r}
names(fang)
names(snp_position)
```

'fang' has 986 columns and 'snp_position' has 15 columns. This information is also available in the environment tab.

The following provides more information than names() does.
```{r}
str(fang)
str(snp_position)
```

This command displays the levels of each variable. For example, the Sample_ID column is a factor with 2782 levels, which makes sense because these should all be unique. Group has 16 levels. Interestingly, in the snp_position, Position is a factor, and not a numeric data type.

## Data Processing

First, I need to transpose the fang file. 

```{r}
fang_transposed <- t(fang)
```

The output has renamed all of the columns as V1, V2, V3, etc. This isn't really a problem, except now I will be searching within the Group row in order to find the maize and teosinte genotypes.It might be easier to transpose the results after filtering the data based on teosinte or maize groups first.

```{r}
maize <- filter(fang, Group %in% c('ZMMMR','ZMMIL', 'ZMMLR'))
teosinte <- filter(fang, Group %in% c('ZMPBA', 'ZMPIL', 'ZMPJA'))
```

Transposing both of these data frames results in a matrix.

```{r}
maize_t <- t(maize)
teosinte_t <- t(teosinte)
```

The first three rows of each transposed matrix need to be removed before joining with the SNP Position data frame.

```{r}
teosinte_t <- teosinte_t[-1:-3,]
maize_t <- maize_t[-1:-3,]
```

I next only selected the relevant columns from SNP_Position.txt.

```{r}
snp_working <- select(snp_position, SNP_ID, Chromosome, Position)
```

I assigned the values in the first column of snp_working to the row names.

```{r}
rownames(snp_working) <- snp_working[,1]
snp_working[,1] <- NULL
```

Now I will join snp_working and the transposed maize and teosinte files.

```{r}
teosinte_t <- as.data.frame(teosinte_t)             #first make sure that the transposed matrices behave like data frames
maize_t <- as.data.frame(maize_t)
maize_final <- merge(snp_working, maize_t, by="row.names")
teosinte_final <- merge(snp_working, teosinte_t, by="row.names")
```


Now I am ready to sort the final data frames and create the output files.

```{r}
setwd("../abby_546X")

```

```{r}
setwd("../abby_546X/R_assignment")
```



```{unix}
cd ..
cd R_Assignment
mkdir maize
mkdir teosinte
```

```{r}
setwd("maize")
```

```{r}
maize_final %>% arrange(Chromosome, Position) %>% filter(Chromosome == 9) %>% write.table("maize_chr9_for.txt", sep="\t",col.names = FALSE, row.names = FALSE)
```

```{r}
maize_final %>% arrange(Chromosome, desc(Position)) %>% filter(Chromosome == 10) %>% write.table("maize_chr1_rev.txt", sep="\t",col.names = FALSE, row.names = FALSE)
```


Again, the working directory of the terminal must also be set to the maize folder.


```{r}
setwd("teosinte")
```

```{r}
teosinte_final %>% arrange(Chromosome, Position) %>% filter(Chromosome == 9) %>% write.table("teosinte_chr9_for.txt", sep="\t",col.names = FALSE, row.names = FALSE)
```

Now I will replace the '?' values with '-' instead.

```{r}
teosinte2 <- data.frame(lapply(teosinte_final, as.character), stringsAsFactors=FALSE)
teosinte2 <- sapply(teosinte2,function(x) {x <- gsub("?","-",x,fixed=TRUE)})
teosinte2 <- as.data.frame(teosinte2)
teosinte2 %>% arrange(Chromosome, desc(Position)) %>% filter(Chromosome == 9) %>% write.table("teosinte_chr9_rev.txt", sep="\t",col.names = FALSE, row.names = FALSE)

```

```{r}
maize2 <- data.frame(lapply(maize_final, as.character), stringsAsFactors=FALSE)
maize2 <- sapply(maize2,function(x) {x <- gsub("?","-",x,fixed=TRUE)})
maize2 <- as.data.frame(maize2)
maize2 %>% arrange(Chromosome, desc(Position)) %>% filter(Chromosome == 9) %>% write.table("maize_chr9_rev.txt", sep="\t",col.names = FALSE, row.names = FALSE)
```

## Data Visualization

```{r}
library(reshape2)
melt(maize_final)
melt(teosinte_final)
```


## Including Plots

I need to combine both teosinte and maize. (Homework instructions were vague.)
```{r}
plot_genotypes <- filter(fang, Group %in% c('ZMMMR','ZMMIL', 'ZMMLR', 'ZMPBA', 'ZMPIL', 'ZMPJA'))
ggplot(data = plot_genotypes) + geom_bar(mapping=aes(x=Group)) #This is my plot of choice.



```

```{r}
t_plot_genotypes <- t(plot_genotypes)
t_plot_genotypes <- t_plot_genotypes[-1:-3,]
plot_gen_final <- merge(snp_working, t_plot_genotypes, by="row.names")

ggplot(data = plot_gen_final) + geom_bar(mapping=aes(x=Chromosome))

```

```{r}
ggplot(data = plot_gen_final) + geom_density(mapping=aes(x=Chromosome))

```
```{r}
test_gen_final <- plot_gen_final
```



```{r}
newColNameIni = "B"
iter = 0 #Initialized an iterator I ended up not using
tempCol = c(NULL) #Initialize new percentage column
tempCol2 = c(NULL) #Initialize missing value column

for (col in colnames(test_gen_final)) { #Iterates through columns of dataframe
  iter <- iter + 1 #The unused iterator iterating
  tempName <- as.name(paste(c(newColNameIni,toString(iter)),collapse="")) #Naming new column
  tempCol = c(NULL) #Initialize column to be added
  for (row in rownames(test_gen_final)) { #Iterates through rows of tDatFram dataframe
    tempString <- toString(test_gen_final[row,col]) #Converts tDatFrame values to true string
    tempVal <- "Heterozygous" #Assumes heterozygous
    if (substring(tempString,1,1) == "?") { #If loop comparing SNP nts
      tempVal <- "NA" #Changes to NA if missing value
    }
    else if (substring(tempString,1,1) == substring(tempString,3,3)) { #If loop comparing SNP nts
      tempVal <- "Homozygous" #Changes to homozygous if if loop true
    }
    tempCol = c(tempCol,tempVal) #Adds to the column to be added
  }
  test_gen_final[[tempName]] <- tempCol #Adds column with new name
}

##Determining the percent homozygous
tempCol = c(NULL) #Initialize new percentage column
tempCol2 = c(NULL) #Initialize missing value column
for (row in rownames(test_gen_final)) { #Iterate through dataframe rows
  countTot = 0 #Initialize total count
  countHom = 0 #Initialize homozygous count
  countNA = 0 #Initialize missing count
  for (iterB in 1:iter) { #Iterate through the columns added previously
    countTot <- countTot + 1 #Always increment total counter
    if (test_gen_final[row,paste(c(newColNameIni,toString(iterB)),collapse="")]=="Homozygous") { #Ask if homozygous
      countHom <- countHom + 1 #Increment homozygous count if homozygous
    }
    else if (test_gen_final[row,paste(c(newColNameIni,toString(iterB)),collapse="")]=="NA") { #Ask if missing
      countNA <- countNA + 1 #Increment missing count if missing
    }
  }
  tempCol <-c(tempCol, countHom/countTot) #Save ratio to percentage column
  tempCol2 <-c(tempCol2, countNA/countTot) #Save missing percentage to a new column
}
test_gen_final$Percent <- tempCol #Add percentage column
test_gen_final$Missing <- tempCol #Add missing data column



```



```{r}
##Graph
barplot(test_gen_final$Percent, names.arg = rownames(test_gen_final), xlab = "site", ylab = "Fraction Homozygous Alleles")
```



```{r}
barplot(test_gen_final$Missing, names.arg = rownames(test_gen_final), xlab = "site", ylab = "Fraction Missing Data")
```




